### 2020-04-15 probabilisticProgrammingMartingales.tex
Tidying up comments

The one-step reduction, \changed[lo]{$\to$}, is given by the function $\red : \Lambda \times S \to \Lambda \times S$ where
\begin{equation}
\red(M,s) = \left\{
    \begin{array}{ll}
        (E[N],s) & \text{if } M = E[R], R \to N \text{ and } R \neq \tsample \\
        (E[\underline{\pi_h(s)}],\pi_t(s)) & \text{if } M = E[\tsample] \\
        (M,s) & \text{if } M \text{ normal form}
    \end{array} \right .
\end{equation}

@@

\lo{Your operational semantics does not maintain a record of the current weight of the reduction.
Note that a.s.~termination does depend on $\tscore$: see \cite[\S 4.3]{DBLP:journals/corr/abs-2004-03924}\footnote{\url{https://arxiv.org/abs/2004.03924}}.}

@@

    \akr{I'm not sure whether $Rch(M)$ is a measurable subset of $\Lambda$, so it might not have a natural measure, but I'm pretty sure this doesn't actually matter, and the $\sigma$-algebra at least is well-defined. $Rch(M)$ could fail to be measurable if there are measurable functions $\mathbb R^n \to \mathbb R$ with non-measurable ranges. I don't know whether these exist. They do for the Lebesgue $\sigma$-algebra, but I'm using the Borel ones.}
    \lo{For the integral to be well-defined, we need $x \mapsto E[\underline x]$ to be a measurable function $\mathbb{R} \to \mathit{Rch}(M)$ (I am sure it is, assuming an appropriate definition of SPCF terms as a measurable space.).}

@@

Given $M'_n$, there is some finite bound on the number of reduction steps that can take place from $M'_n$ without a $\tY$-reduction step, because of the type system, therefore $y_{n+1}$ is (conditional on $\mathcal{F}_{y_n+1}$) a bounded stopping time, therefore $\mathbb{E}[f(M_{y_{n+1}}) \mid \mathcal{F}_{y_n+1}] \leq f(M_{y_n+1})$. If $M_{y_n}$ isn't already in normal form, then $M_{y_n} = E[\tY N]$ for some $E, N$, therefore $M_{y_n+1} = N (\tY N)$ and $f(M_{y_n+1}) \leq f(M_{y_n}) - 1$, therefore if $M_{y_n}$ isn't normal form, $\mathbb{E}[X_{n+1} \mid \mathcal{F}_{y_n+1}] \leq X_n - 1$.

Overall, this implies that $\mathbb{E}[X_{n+1}] - \mathbb{E}[X_{n}] \leq -\mathbb P[M_{y_n} \text{not n.f.}]$ therefore as $\mathbb{E}[X_n]$ is bounded below, $\mathbb P[M_{y_n} \text{not n.f.}]$ must tend to 0 as $n \to \infty$. $M_{y_n} \text{n.f.} \Rightarrow M_{y_{n+1}} \text{n.f.}$ therefore 
\begin{align*}
\mathbb P[M_{y_n} \text{not n.f.~for infinitely many values of $n$}]  
& = \mathbb P[\forall n: M_{y_n} \text{not n.f.}] \\
& \leq \inf_n \mathbb P[M_{y_n} \text{not n.f.}] \\
& = 0
\end{align*} 
therefore $M$ terminates almost surely.
\end{proof}

>>>>

The one-step reduction is given by the function $\red : \Lambda \times S \to \Lambda \times S$ where
\begin{equation}
\red(M,s) = \left\{
    \begin{array}{ll}
        (E[N],s) & \text{if } M = E[R], R \to N \text{ and } R \neq \tsample \\
        (E[\underline{\pi_h(s)}],\pi_t(s)) & \text{if } M = E[\tsample] \\
        (M,s) & \text{if } M \text{ normal form}
    \end{array} \right .
\end{equation}
\lo{Replace (globally) ``normal form'' by ``value''.}

@@

\lo{Your operational semantics does not maintain a record of the current weight of the reduction.
A.s.~termination does depend on $\tscore$: see \cite[\S 4.3]{DBLP:journals/corr/abs-2004-03924}\footnote{\url{https://arxiv.org/abs/2004.03924}}.
I think it important to take the behaviour of $\tscore$ into account;
you should do it as a future task.}

@@

    \akr{I'm not sure whether $Rch(M)$ is a measurable subset of $\Lambda$, so it might not have a natural measure, but I'm pretty sure this doesn't actually matter, and the $\sigma$-algebra at least is well-defined. $Rch(M)$ could fail to be measurable if there are measurable functions $\mathbb R^n \to \mathbb R$ with non-measurable ranges. I don't know whether these exist. They do for the Lebesgue $\sigma$-algebra, but I'm using the Borel ones.}
    \lo{It is worth inserting the preceding (including the example of a measurable function with non-measurable range) as a remark (or as a ``commenting.sty'' note).}
    \lo{For the integral to be well-defined, we need $x \mapsto E[\underline x]$ to be a measurable function $\mathbb{R} \to \mathit{Rch}(M)$.}

@@

Given $M'_n$, there is some finite bound on the number of reduction steps that can take place from $M'_n$ without a $\tY$-reduction step, because of the type system, therefore $y_{n+1}$ is (conditional on $\mathcal{F}_{y_n+1}$) a bounded stopping time, therefore $\mathbb{E}[f(M_{y_{n+1}}) \mid \mathcal{F}_{y_n+1}] \leq f(M_{y_n+1})$. 
If $M_{y_n}$ isn't already in normal form, then \changed[lo]{$M_{y_n} = E[\tY \lambda x.N]$ for some $E, N$, therefore $M_{y_n+1} = E[\lambda z. N[(\tY \lambda x. N)/x] z]$} and $f(M_{y_n+1}) \leq f(M_{y_n}) - 1$, therefore if $M_{y_n}$ isn't normal form, $\mathbb{E}[X_{n+1} \mid \mathcal{F}_{y_n+1}] \leq X_n - 1$.

Overall, this implies \lo{This implication is still not clear to me.} that $\mathbb{E}[X_{n+1}] - \mathbb{E}[X_{n}] \leq -\mathbb P[M_{y_n} \text{not n.f.}]$ therefore as $\mathbb{E}[X_n]$ is bounded below, $\mathbb P[M_{y_n} \text{not n.f.}]$ must tend to 0 as $n \to \infty$. $M_{y_n} \text{n.f.} \Rightarrow M_{y_{n+1}} \text{n.f.}$ therefore 
\begin{align*}
\mathbb P[M_{y_n} \text{not n.f.~for infinitely many values of $n$}]  
& = \mathbb P[\forall n: M_{y_n} \text{not n.f.}] \\
& \leq \inf_n \mathbb P[M_{y_n} \text{not n.f.}] \\
& = 0
\end{align*} 
therefore $M$ terminates almost surely.
\end{proof}
